---
title: "Assignment 2"
author: "Liam McGrath"
format:
  html:
    toc: true
    toc-location: left
    self-contained: true
    theme: "vapor"
---

```{r}
library(tidyverse)
library(caret)
library(rpart)
library(rpart.plot)
library(randomForest)
library(e1071)
library(caret)
library(pROC)
library(performanceEstimation)
```




# PART I: Collect, Explore, and Prepare the Data

1. Import and preview the data. Call the dataset loans. Remember to use the
tidyverse package as much as possible. This includes making use of the
read_csv() functions when importing data. Use the glimpse() function
to show that your data was imported properly.

```{r}
loans <- read_csv("https://s3.amazonaws.com/notredame.analytics.data/lendingclub.csv")

glimpse(loans)
```

2. You notice that the Grade, EmploymentLength, HomeOwnership,
IncomeVerified, LoanPurpose, and Default features were
imported as string values. Convert these features to factors. Use the
glimpse() function to show that your conversion worked.


```{r}
loans <- loans %>%
  mutate(Grade = as.factor(Grade),
         EmploymentLength = as.factor(EmploymentLength),
         HomeOwnership = as.factor(HomeOwnership),
         IncomeVerified = as.factor(IncomeVerified),
         LoanPurpose = as.factor(LoanPurpose),
         Default = as.factor(Default))

glimpse(loans)
```

3. Get descriptive statistics for the entire dataset.

```{r}
summary(loans)
```


4. Split your dataset into training and test sets using stratified random sampling
(with Default as the stratum). Assign 75% of your original data to the
training set and 25% to the test set. Call the training set, loans_train,
and the test set, loans_test. Use ‘1234’ for your seed and don’t forget to
use the RNGkind() function.

```{r}
set.seed(1234)
RNGkind("Mersenne-Twister")

inTrain <- createDataPartition(y = loans$Default, p = 0.75, list = FALSE)

loans_train <- loans[inTrain, ]
loans_test <- loans[-inTrain, ]
```


5. Based on the descriptive statistics of the loans dataset, you notice that the
dataset suffers from class imbalance. Use the smote() function from the
performanceEstimation package to balance (50-50) the training set.
Use ‘1234’ for your seed here as well and don’t forget to use the
RNGkind() function

```{r}
set.seed(1234)

RNGkind("Mersenne-Twister")

loans_train_balanced <- smote(Default ~ ., data = loans_train, perc.over = 100, k = 5)

```



6. During the recursive partitioning process, decision trees attempt to create a
split using different values for a feature. You notice that the values for the
Delinquencies, Inquiries, OpenAccounts, TotalAccounts,
and PublicRecords features are double-precision values. This means that
the decision tree could split on OpenAccounts <= 1.4. This makes no
sense since a borrower cannot have 1.4 open accounts. To resolve this
potential issue, convert these features to integers for both the training and test
sets. Use the glimpse() function to show that your conversion worked.

```{r}
loans_train_balanced <- loans_train_balanced %>%
  mutate(Delinquencies = as.integer(Delinquencies),
         Inquiries = as.integer(Inquiries),
         OpenAccounts = as.integer(OpenAccounts),
         TotalAccounts = as.integer(TotalAccounts),
         PublicRecords = as.integer(PublicRecords))

glimpse(loans_train_balanced)
```



# PART II: Train the Model

1. Train a model using the rpart() function from the rpart package. Set the
complexity parameter to 0.005

```{r}
loans_tree <- rpart(Default ~ ., data = loans_train_balanced, method = "class", cp = 0.005)
```


2. Create a visualization of the decision tree model.

```{r}
rpart.plot(loans_tree)
```


3. Based on the visualization, which feature is most predictive of whether a borrower will default or not default on their loan?

based on the visualization, the most predictive feature of whether a borrower will default or not default on their loan is the grade of the loan. 


 


# PART III: Evaluate the Model

1. Using the model that you created in the previous section, predict the outcomes
of the loans in the test dataset. Preview the first 5 predictions

```{r}
loans_test$Prediction <- predict(loans_tree, newdata = loans_test, type = "class")


head(loans_test$Prediction, 5)
```


2. Add a new column to the loans_test dataset with the predicted outcome
for each loan. Call the new column Prediction. Use the glimpse()
function to show that the column was added.

```{r}

loans_test$Prediction <- predict(loans_tree, newdata = loans_test, type = "class")

glimpse(loans_test)

```


3. Calculate the accuracy of your model against the test data

```{r}
confusionMatrix(loans_test$Prediction, loans_test$Default)$overall["Accuracy"]

```


# PART IV: Interpret the Results
Assuming that the Prediction column in loans_test is what your model
predicts will happen, while the Default column is what actually happened, answer
the following questions:

1. On average, for every loan that is paid back in full, LendingClub earns 12% of the original loan amount as profit. With this information, how much money did LendingClub make on the loans that were issued and paid back in full?

```{r}
loans_test %>%
  filter(Prediction == "No", Default == "No") %>%
  summarise(TotalProfit = sum(LoanAmount * 0.12))
```
based on the code above, LendingClub made $10814190 on the loans that were issued and paid back in full.




2. On average, LendingClub loses 70% of the original loan amount, for everyloan that was issued and not paid back in full. How much money did LendingClub lose as a result of bad loans that were issued based on the recommendation of your model?

```{r}

loans_test %>%
  filter(Prediction == "Yes", Default == "No") %>%
  summarise(TotalLoss = sum(LoanAmount * 0.70))


```
on average, LendingClub lost 6008923 on the loans that were issued and not paid back in full.

3. Assuming that LendingClub incurs a loan servicing cost of 2% (of the original loan amount) for every loan that it issues, did LendingClub realize a profit or a loss based on your model? Make sure to show your work as code.


```{r}



loans_test <- loans_test %>%
  mutate(ServicingCost = LoanAmount * 0.02)

summary_df <- loans_test %>%
  group_by(Prediction) %>%
  summarise(TotalServicingCost = sum(ServicingCost),
            TotalProfit = sum(LoanAmount * 0.12),
            TotalLoss = sum(LoanAmount * 0.70)) %>%
  mutate(NetProfit = TotalProfit - TotalLoss - TotalServicingCost)

print(summary_df)


```

based on the code above, LendingClub realized a profit of $12707865 based on the model.




4. What about the opportunity cost? How much money would LendingClub
have made from loans that your model suggested were bad but were actually
good?

```{r}


opportunity_cost <- loans_test %>%
  filter(Prediction == "No", Default == "No") %>%
  mutate(PotentialProfit = LoanAmount * 0.12) %>%  # Assuming 12% 
  summarise(TotalOpportunityCost = sum(PotentialProfit))

print(opportunity_cost)






```
the opportunity cost for LendingClub would have been $10814190 based on the model.

---
