---
title: "Assignment 1"
author: "Liam McGrath"
format: html
editor: visual
theme: "vapor"
---

Load Packages

```{r}
library(tidyr)
library(ggplot2)
library(dplyr)
library(readr)
library(stringr)
```

##PART I: Collect and Describe the Data

Load the inspections dataset

```{r}

inspections <- read_csv("https://s3.amazonaws.com/notredame.analytics.data/inspections.csv")

glimpse(inspections)



```

Changing names and data types of the data in inspections

```{r}

inspections <- inspections %>%
  rename(ID = `Inspection ID`, DBANAME = `DBA Name`, AKANAME = `AKA Name`, License = `License #`,
         InspectionDate = `Inspection Date`,InspectionType = `Inspection Type`, 
         FacilityType = `Facility Type`, Risk = Risk, Address = Address, City = City, State = State) %>%
  mutate(ID = as.integer(ID), DBANAME = as.character(DBANAME), AKANAME = as.character(AKANAME), License = as.character(License), FacilityType = as.factor(FacilityType), Risk = as.factor(Risk), Address = as.character(Address), City = as.factor(City), State = as.factor(State),Zip = as.factor(Zip),Results = as.factor(Results), InspectionDate = as.character(InspectionDate), InspectionType = as.factor(InspectionType), Latitude = as.double(Latitude), Longitude = as.double(Longitude), Location = as.character(Location))
         


inspections$InspectionDate <- as.Date(inspections$InspectionDate, format = "%m/%d/%Y")

glimpse(inspections)


summary(inspections$City)

summary(inspections$State)

summary(inspections$Zip)

summary(inspections$InspectionDate)

summary(inspections$Longitude)

summary(inspections$Latitude)
```

##PART II: Resolve Data Quality Issues

Question: 1

```{r}
inspections$Zip[inspections$Zip == "60627"] <- "60827"

summary(inspections$Zip)
```

Question: 2

```{r}
inspections$City[inspections$City == "BRIDEVIEW"] <- "Bridgeview"

inspections$Zip[inspections$DBANAME == "STEVES 77 ICE CREAM"] <- "60455"

summary(inspections$City)

```

Question: 3

```{r}

inspections$DBANAME[is.na(inspections$Zip)]

inspections$Zip[inspections$DBANAME == "CHARTWELLS"] <- "60455"

inspections$Zip[inspections$DBANAME == "DUNKIN DONUTS"] <- "60455"

```

Question 4

```{r}
zipcodes <- read_csv("C:/Users//liamm/Downloads/zipcode.csv")

glimpse(zipcodes)
```



```{r}


inspections <- left_join(inspections, zipcodes, by = c("Zip" = "zip"))

glimpse(inspections)



inspections <- inspections %>%
  select(-City, -State, -Latitude, -Longitude)



inspections <- inspections %>%
  mutate(city = as.factor(city), state = as.factor(state), Zip = as.factor(Zip), latitude = as.double(latitude), longitude = as.double(longitude))

glimpse(inspections)
```


Question 6
```{r}

summary(inspections$city)

summary(inspections$state)

summary(inspections$Zip)

summary(inspections$longitude)

summary(inspections$latitude)
```
##PART III: Visualize and Interpret the Data


Question 1

```{r}
inspections <- inspections %>%
  mutate(ViolationCount = str_count(Violations, "\\|") + 1)

```

Question 2

```{r}
ggplot(inspections, aes(x = Results, y = ViolationCount)) +
  geom_boxplot() +
  labs(title = "Relationship between Inspection Violations and Inspection Results",
       x = "Results",
       y = "Violation Count")
```


The boxplot represnts the relationship between the number of violations and the 
inspection results. As shown in the chart above, the number of violations seems to
have an effect on the inspection results. The more violations that are present the lower the chance of passing is. Having a violation count above 5 typically means that you will fail your inspection.


